---
title: "Sargasso arribazon"
subtitle: "Some graphs for Nancy"
date: "`r Sys.Date()`"
author: 
  - name: Diego Lizcano
    orcid: https://orcid.org/0000-0002-9648-0576
  - name: Nancy Cabanillas-Terán
    orcid: https://orcid.org/0000-0003-1268-5900
license: CC BY-SA
toc: true
format: 
  html:
    theme: 
      light: flatly
      dark: darkly
    code-fold: true
    code-block-bg: true
    code-block-border-left: "#31BAE9"
citation: true
google-scholar: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  dev = "png",
  dev.args = list(type = "cairo-png"),
  fig.width = 7,
  fig.height = 5,
  fig.align = "center",
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  error = FALSE,
  message = FALSE,
  cache=TRUE)
```


## Question

Which is the relationship distance to coast line and DO?

## Set up analysis

Load libraries and set some options.

```{r set_up}
#| warning: false
#| message: false

library(readxl)
library(gt)
library(lubridate)
library(stringr)
library(readxl)
library(sf)
# library(MuMIn) # multimodel inference
library(visreg) # see trend
library(MASS) # stepAIC
# library(terra)
library(lme4) # mix model
library(sjPlot)
library(see)
library(performance)
library(mapview)
library(corrplot)
library(DT)
library(measurements) # transform DMS to lat lon
library(ggeffects)

library(tidyverse)



options(scipen=99999)
options(max.print=99999)
options(stringsAsFactors=F)
```

## Load and Map the data

```{r}
sargasso <- read_excel("C:/CodigoR/Nancy2/data/parametros_fisicoquimicos2023.xlsx", 
    sheet = "alta_baja")


sargasso$lat <- as.numeric(conv_unit(sargasso$Lattitude, from = "deg_min_sec", to = "dec_deg"))
sargasso$lon <- conv_unit(sargasso$Longitude , from = "deg_min_sec", to = "dec_deg") 
sargasso$lon <- as.numeric(sargasso$lon)*(-1)

# convierte covs a puntos terra
puntos <- vect(sargasso, geom=c("lon", "lat"), crs="EPSG:4326")
# convierte a sf
sargasso_sf <- sf::st_as_sf(puntos)

mapview(sargasso_sf, zcol="DO_mgL")

```

## Relation DO[mgL] and distance to coastline, pluma, temporada, and locality

### ggplot graph (Loess method)

```{r}
# Add the regression line Loess method
ggplot(sargasso, aes(x=Distance_coast, y=DO_mgL, color=Pluma)) + 
  geom_point()+
  geom_smooth() + # method=lm
  facet_grid(Temporada ~ Locality, scales = "free")

# # Marginal density plot of x (top panel)
# xdensity <- ggplot(sargasso, aes(Distance_coast, fill=Pluma)) + 
#   geom_density(alpha=.5) + 
#   scale_fill_manual(values = c('#999999','#E69F00')) + # + 
#   # theme(legend.position = "none")
#    facet_grid(Temporada ~ Locality, scales = "free")
# xdensity

```


### Regresion models

#### A simple model

$$
DO_mgL \sim DistanceCoast + Pluma + Temporada
$$

```{r}
lm1 <- lm(DO_mgL ~ Distance_coast + Pluma + Temporada, data = sargasso)


dat <- predict_response(lm1, terms = c("Distance_coast", "Pluma", "Temporada"))
plot(dat, facets = TRUE)

out1 <- check_model(lm1)
plot(out1)

summary(lm1)
tab_model(summary(lm1))
#visreg(glm2, xvar = c("Distance_coast"), 
#       yvar="DO_mgL",
#       by="Pluma")

```


#### A model with 2 interaction between distance and temporada

$$
DO_mgL \sim DistanceCoast * Temporada + Pluma
$$

```{r}
lm2 <- lm(formula = DO_mgL ~ Distance_coast * Temporada + Pluma ,   data = sargasso)

dat2 <- predict_response(lm2, terms = c("Distance_coast", "Pluma", "Temporada"))
plot(dat2, facets = TRUE)

out2 <- check_model(lm2)
plot(out2)

tab_model(summary(lm2))

```
 
 complex to interpret 


#### Three-Way-Interaction

$$
DO_mgL \sim DistanceCoast * Temporada * Pluma
$$

```{r}
lm3 <- lm(formula = DO_mgL ~ Distance_coast * Temporada * Pluma ,   data = sargasso)

plot_model(lm3, type = "int")

out3 <- check_model(lm3)
plot(out3)

tab_model(summary(lm3))
```


## Información de la sesión en R.

```{r sesion, results='markup'}
sessionInfo()
```




