---
title: "Corales reclutamiento"
author: "Diego L. Lizcano"
date: "16 May 2017"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Coral and recruitment

The main questions to be addressed in this investigation are: 

1. Are there any coral recruits coming through the channels to colonize the bay?.
    
2. How high is the abundance level of coral recruitment in the reef lagoon, the channels and the bay?.

3. Are the natural and the artificial channel act differently for coral recruitment?.  

4. Further we want to know the differences in salinity and temperature in the channels, the
lagoon compared to the bay as well as water exchange and near surface water
currents close to the channels to get an idea of the local water movement in the area.


# Study area

“SE Chetumal Bay” (BAY) is surrounded by cays and channels connecting it to the rest of the bay.

The artificial channel (AC), called “Zaragoza Channel” is a man made channel built in the 19 th century, 
closed because of sedimentation and has been reopened from 1996-2000 by the Mexican army.

The natural channel (NC), called “Bacalar Chico” is a historical old, meandered, mangrove channel.

The reef lagoon (RL) which influences the Chetumal Bay through the channels is part of the national park “Xcalak Reefs”.

# Answering the questions

The main questions to be addressed in this investigation are: 

1. Are there any coral recruits coming through the channels to colonize the bay?.
    
    Very hard to answer taking in to account that you are not tracking coral or larvae movement through the channels. We can answer how many recuits are in the bay by genus in a boxplot

2. How high is the abundance level of coral recruitment in the reef lagoon, the channels and the bay?

  We can see the abundance of corals by ploting how many recuits are in the bay by genus in a boxplot (same graph as previous question) 

3. Are the natural and the artificial channel act differently for coral recruitment? 

  We can see if there are statistically significant diferences between Localityes by an ANOVA (1) and also between sites by ANOVA (2). 




## Load the data set and put together tables

```{r read, cache=TRUE, echo=FALSE, message=FALSE}
library(lme4)
library(nlme)
library(arm)
library(sjPlot)
library(dplyr)
library(ggplot2)
library(vegan)

library(sjPlot)
library(glmulti)
library(bbmle)

# read data set
library(readr)
# Qcual cambio a 1
corales <- read.csv("D:/BoxFiles/Box Sync/CodigoR/Nancy2/data/corales3.csv", 
                    sep = ";", header = T)

corales$all <- apply(corales[,35:41], 1, sum) # suma de todos los corales

# tabla de Acropora sp.	Agaricia sp.	Faviidae	Porites spp.	Siderastrea sp Unknown_Coral
corales_algunos1 <- cbind(corales$Acropora.sp. , corales$Agaricia.sp. ,
                          corales$Faviidae , corales$Porites.spp. , 
                          corales$Siderastrea.sp. , corales$Unknown.Coral..too.small.) 

# here in the same comlumn
corales_algunos2 <- as.data.frame(rbind(cbind(corales$Acropora.sp. ,   "Acropora"),
                          cbind(corales$Agaricia.sp. ,   "Agaricia"),
                          cbind(corales$Faviidae ,       "Faviidae"),
                          cbind(corales$Porites.spp. ,   "Porites" ),
                          cbind(corales$Siderastrea.sp. ,"Siderastrea"), 
                          cbind(corales$Unknown.Coral..too.small., "Unknown_Coral") ))

corales_algunos2$V1 <- as.numeric(levels(corales_algunos2$V1)[corales_algunos2$V1]) # convert from factor to numeric

corales_algunos3 <- cbind(corales_algunos2, c(as.character(rep(corales$Locality, 6))), 
                                             c(as.character(rep(corales$Site, 6))) )                       
colnames(corales_algunos3) <- c("recruitment", "genus", "Locality", "site")
corales_algunos3 <-  as.data.frame(corales_algunos3) # make data frame


# suma de las abundancias de los corales seleccionados
corales$corales_algunos_sum <- apply(corales_algunos1, 1, sum) 

head(corales)

```



# BoxPlot

Let see how recruitment varies per site

```{r, boxplot1, echo=TRUE, cache=TRUE, fig.cap="Sumarazing how many recuits are in the bay first agregating all corals as a sum."}
tabla2 <- as.data.frame(cbind(as.character(corales[,2]), as.numeric( corales[,43])))
colnames(tabla2) <- c("site", "abund")
tabla2$abund <- as.numeric(tabla2$abund)
tabla3 <- table(tabla2) # abundancia de selected corales and locality
tabla4 <- apply(tabla3, 1, mean) # suma selected corales by locality
# barplot(tabla4)

ggplot(tabla2, aes(x = site, y = abund)) +
  geom_boxplot(fill = "grey80", colour = "blue") +
  scale_x_discrete() + xlab("site") +
  ylab("Recruitment (sum) of several coral genus")


tabla_sitio <- corales %>% 
     group_by(Site) %>% 
     summarise (Amphipoda = as.numeric(sum(Amphipoda)), 
                Anemone= as.numeric(sum(Anemone)),
                Ascidiacea = as.numeric(sum(Ascidiacea)),
                Bivalvia= as.numeric(sum(Bivalvia)),
                Branching_coralline_algae =as.numeric(sum(Branching.coralline.algae)),
                Branching_erect_algae= as.numeric(sum(Branching.erect.algae)),
                Bryozoa = as.numeric(sum(Bryozoa)),
                Coarsely.branched.algae= as.numeric(sum(Coarsely.branched.algae)),
                Crustacea= as.numeric(sum(Crustacea)),
                Crustose.algae=as.numeric(sum(Crustose.algae)),
                Cyanophycea= as.numeric(sum(Cyanophycea)),
                Echinoidea= as.numeric(sum(Echinoidea)),
                Filamentous.algae= as.numeric(sum(Filamentous.algae)),
                Foraminifera= as.numeric(sum(Foraminifera)),
                Holothuroidea= as.numeric(sum(Holothuroidea)),
                Hydrozoa= as.numeric(sum(Hydrozoa)),
                Jointed.calcareus.algae= as.numeric(sum(Jointed.calcareus.algae)),
                Mollusca= as.numeric(sum(Mollusca)),
                Patellidae=as.numeric(sum(Patellidae)),
                Polychaeta= as.numeric(sum(Polychaeta)),
                Porifera=as.numeric(sum(Porifera)),
                Sheet.like.algae= as.numeric(sum(Sheet.like.algae)),
                Sipunculidae=as.numeric(sum(Sipunculidae)),
                Thick.leathery.alga=as.numeric(sum(Thick.leathery.algae)),
                turf.algae= as.numeric(sum(turf.algae)),
                unbranched.erect.algae = as.numeric(sum(unbranched.erect.algae)),
                Vermetidae=as.numeric(sum(Vermetidae)),
                Zoanthidae = as.numeric(sum(Zoanthidae)),
                Acropora.sp.=as.numeric(sum(Acropora.sp.)),
                Agaricia.sp.=as.numeric(sum(Agaricia.sp.)),
                Faviidae=as.numeric(sum(Faviidae)),
                Porites.spp.=as.numeric(sum(Porites.spp.)),
                Siderastrea.sp.=as.numeric(sum(Siderastrea.sp.)),
                Unknown.Coral..broken.=as.numeric(sum(Unknown.Coral..broken.)),
                Unknown.Coral..too.small.=as.numeric(sum(Unknown.Coral..too.small.)))

tabla_sitio$suma <- apply(tabla_sitio[,2:32], 1, sum)
tabla_sitio[,c(1,37)]


tabla_Locality <- corales %>% 
     group_by(Locality) %>% 
     summarise (Amphipoda = as.numeric(sum(Amphipoda)), 
                Anemone= as.numeric(sum(Anemone)),
                Ascidiacea = as.numeric(sum(Ascidiacea)),
                Bivalvia= as.numeric(sum(Bivalvia)),
                Branching_coralline_algae =as.numeric(sum(Branching.coralline.algae)),
                Branching_erect_algae= as.numeric(sum(Branching.erect.algae)),
                Bryozoa = as.numeric(sum(Bryozoa)),
                Coarsely.branched.algae= as.numeric(sum(Coarsely.branched.algae)),
                Crustacea= as.numeric(sum(Crustacea)),
                Crustose.algae=as.numeric(sum(Crustose.algae)),
                Cyanophycea= as.numeric(sum(Cyanophycea)),
                Echinoidea= as.numeric(sum(Echinoidea)),
                Filamentous.algae= as.numeric(sum(Filamentous.algae)),
                Foraminifera= as.numeric(sum(Foraminifera)),
                Holothuroidea= as.numeric(sum(Holothuroidea)),
                Hydrozoa= as.numeric(sum(Hydrozoa)),
                Jointed.calcareus.algae= as.numeric(sum(Jointed.calcareus.algae)),
                Mollusca= as.numeric(sum(Mollusca)),
                Patellidae=as.numeric(sum(Patellidae)),
                Polychaeta= as.numeric(sum(Polychaeta)),
                Porifera=as.numeric(sum(Porifera)),
                Sheet.like.algae= as.numeric(sum(Sheet.like.algae)),
                Sipunculidae=as.numeric(sum(Sipunculidae)),
                Thick.leathery.alga=as.numeric(sum(Thick.leathery.algae)),
                turf.algae= as.numeric(sum(turf.algae)),
                unbranched.erect.algae = as.numeric(sum(unbranched.erect.algae)),
                Vermetidae=as.numeric(sum(Vermetidae)),
                Zoanthidae = as.numeric(sum(Zoanthidae)),
                Acropora.sp.=as.numeric(sum(Acropora.sp.)),
                Agaricia.sp.=as.numeric(sum(Agaricia.sp.)),
                Faviidae=as.numeric(sum(Faviidae)),
                Porites.spp.=as.numeric(sum(Porites.spp.)),
                Siderastrea.sp.=as.numeric(sum(Siderastrea.sp.)),
                Unknown.Coral..broken.=as.numeric(sum(Unknown.Coral..broken.)),
                Unknown.Coral..too.small.=as.numeric(sum(Unknown.Coral..too.small.)))



```



## Testing diferences by ANOVA

### diferences in coral recruitment by a piosson procces determined by genus

```{r glm01, echo=TRUE, cache=TRUE, fig.cap="The graph shows the multiple comparison (Tukey test) between Sites.  The pairs having p adj lower to 0.05 are differents.", fig.height=9}

glm2 <- glm(formula = recruitment ~ genus, family = poisson(),data = corales_algunos3)
model2 <- aov(glm2)
summary(model2)

# glm3 <- glm (corales_algunos_sum ~  Site_name, family = poisson(), data = corales)

# summary(glm1)
# summary(glm2)
# summary(glm3)
# anova(glm3, test = "Chisq")
# anova1 <- aov(glm3)
# (posthoc1 <- TukeyHSD(x=anova1, "Site_name", conf.level=0.95))
# plot(posthoc1)

```



## The next graph shows how coral recruitment varies according to genus

```{r, boxplot2, echo=TRUE, fig.cap="The second boxplot shows recruitment by locality wraped by genus"}

library(ggplot2)

ggplot(corales_algunos3, aes(x = Locality , y = recruitment)) +
  geom_boxplot(fill = "grey80") + facet_wrap(~genus) +
  scale_x_discrete() + xlab("Locality") +
  ylab("Recruitment of selected corals") +
  theme(strip.text.x = element_text(face = "italic"))

```



## Testing diferences in ANOVA

as a piosson procces determined by  site = locality name (factor) 

```{r glm1, echo=TRUE, cache=TRUE, fig.cap="The graph shows the multiple comparison (Tukey test) between Localities. The pairs RL-AC and  NC-BAY are not differents.", fig.height=9}


glm4 <- glm (corales_algunos_sum ~  Locality, family = poisson(), data = corales)

# summary(glm1)
# summary(glm2)
# summary(glm3)
summary(glm4)

anova(glm4, test = "Chisq")
anova1 <- aov(glm4)
(posthoc1 <- TukeyHSD(x=anova1, "Locality", conf.level=0.95))
plot(posthoc1)


```

Las localidades BAY y NC son significativamente diferentes a AC y RL. Y como se aprecia en el boxplot (Fig. 1) tienen menor reclutamiento de coral.

```{r, anova2, echo=TRUE, cache=TRUE, fig.cap="The graph shows the multiple comparison (Tukey test) between sites The pairs Porites-Acropora, Porites-Agaricia, Porites-Faviidae, Siderastrea-Porites and Unknown_Coral-Porites are different.", fig.height=9}
glm5 <- glm (recruitment ~  site + genus,  data = corales_algunos3)
anova(glm5, test = "Chisq")
anova2 <- aov(glm5)
(posthoc2 <- TukeyHSD(x=anova2, "site", conf.level=0.95))
# plot(posthoc2)

```


In most of the comparisons were Porites is included produces a statistically significant diference. As we can see in the Boxplot (Fig. 2) Porites recruitment is bigger than other genus. 




# Predicting recruitmentof all corals as a function of taxa  and  algae richness (several algae). No mater stepwise procedure

```{r pred1}



# algae richness
corales$algae_richnes <- apply(cbind(corales$Branching.coralline.algae , corales$Branching.erect.algae , 
                                    corales$Coarsely.branched.algae , corales$Filamentous.algae , 
                                    corales$Jointed.calcareus.algae , corales$Crustose.algae , 
                                    corales$Sheet.like.algae + corales$Thick.leathery.algae ,
                                    # corales$turf.algae ,
corales$unbranched.erect.algae),1,sum)


glm7 <- glm(corales_algunos_sum ~  Porifera * algae_richnes, family = poisson(), data = corales)

glm9 <- glm(corales_algunos_sum ~  algae_richnes, family = poisson(), data = corales)

glm11 <- glm(corales_algunos_sum ~  Porifera, family = poisson(), data = corales)

glm13 <- glm(corales_algunos_sum ~  Porifera * Filamentous.algae, family = poisson(), data = corales)

glm14 <- glm(corales_algunos_sum ~  Porifera * Crustose.algae, family = poisson(), data = corales)  

glm15 <- glm(corales_algunos_sum ~  Porifera * Thick.leathery.algae, family = poisson(), data = corales)

glm16 <- glm(corales_algunos_sum ~  Porifera * unbranched.erect.algae, family = poisson(), data = corales)

glm17 <- glm(corales_algunos_sum ~  Porifera * Sheet.like.algae, family = poisson(), data = corales)




AICctab(glm7, glm9,  glm11, glm13, glm14, glm15, glm16, glm17, 
        base=T, weights=T, delta=T, sort=T,
    nobs = length(corales))


### Best Model of all!!!!!
summary(glm14)

newdato <- as.data.frame(cbind(Porifera=rep(0:6, each = 10), Crustose.algae=rep(0:1, 10)) )
predichos <- predict(glm14, newdata = newdato, type = "response")

require(geoR)
data(elevation)
library(fields)


grid = expand.grid(list(Porifera = seq(0, 6, 1), Crustose.algae = seq(0, 1, 0.1)))

z = predict(glm14, newdata = grid)
grid$Height = as.numeric(z)

#### with image for predictors with interaction terms#####
psi.matrix <- array(NA, dim = c(100, 100)) # Prediction matrix, for every

# combination of values of two interacting predictors
pr.mat<-data.frame(Porifera=seq(0, 
                                6, 
                                length.out=100),
                   Crustose.algae=seq(0, 1, 
                            length.out=100))

for(i in 1:100){
  for(j in 1:100){
       psi.matrix[i, j]<-predict(glm14, newdata=data.frame(
          Porifera=pr.mat$Porifera[i],
         # mean=pr.mat$mean[j]), 
        #  range=pr.mat$range[j]), 
          Crustose.algae=pr.mat$Crustose.algae[j]),
          type="response")
  }
}

mapPalette <- colorRampPalette(c("blue", "yellow", "orange", "red"))


#plot.par<-par()  #save default graphics parameters

##obtain original unscale values for predictors


# png("BxR_FF.png", width=2000, height=1500, res=300)

#fig<-layout(matrix(c(1:12), 6,2, byrow=F))
par(mar=c(3,5,3,2))
image.plot(x = pr.mat$Porifera, y = pr.mat$Crustose.algae , z = psi.matrix,
           col = mapPalette(100), axes=F,xlab = NA ,ylab = NA, 
           legend.width=1, axis.args=list(cex.axis=0.8, lwd=0,
                              line=-0.5) 
           # ,main="Coral recruitment")
            )
contour(x = pr.mat$Porifera, y = pr.mat$Crustose.algae, z = psi.matrix, 
        add = TRUE, lwd = 1)
axis(side = 1, tck = -.015, labels = NA)
axis(side = 2, tck = -.015, labels = NA)
axis(side = 1, lwd = 0, line = -.6)
axis(side = 2, lwd = 0, line = -.6, las = 1)
mtext(side = 1, "Porifera abundance", line = 1.8)
mtext(side = 2, "Crustose algae", line = 2.3)

points(corales$Porifera, corales$Crustose.algae, pch=19, cex=0.5, 
       col="black")

# dev.off()



########################################################
###### Cross-validation for Generalized Linear Models
#########################################################
library("boot")
cost <- function(r, pi = 0) mean(abs(r - pi) > 0.5)  ## cost function necessary for binomial data
m11.cv <- cv.glm(data = corales, glm14, cost, K = 10)  # use leave-one-out cross validation (can use K-fold cross validation for larger data sets)
## Warning: glm.fit: fitted probabilities numerically 0 or 1 occurred

# Now lets see what our error rate was:

m11.cv$delta
## [1] 0.2381 0.2438
# That’s not too bad.

muhat <- fitted(glm14)
glm14.diag <- glm.diag(glm14)
(cv.err <- mean((glm14$y - muhat)^2/(1 - glm14.diag$h)^2))

glm.diag.plots(glm14, glm14.diag)

#####################################################
## Make your reciever-operater curve
#####################################################

# library(pROC)
# 
# 
# m.roc <- multiclass.roc(corales$corales_algunos_sum, predict(glm8, backtransform = TRUE))
# auc(m.roc)
# ci(m.roc)
# plot(m.roc[[1]], m.roc[[2]])
# m.roc[[7]]

```


# Looking the functional groups 

```{r stepwise}
library(bootStepAIC)
library(lattice)
library(fields)
glm12 <- glm (corales_algunos_sum ~  Porifera + 
                          Branching.coralline.algae + Branching.erect.algae + 
                          Sheet.like.algae +
                          Coarsely.branched.algae + Filamentous.algae + 
                          Jointed.calcareus.algae + Crustose.algae + 
                          Sheet.like.algae + Thick.leathery.algae +
                          # turf.algae + 
                          unbranched.erect.algae , 
                      family = poisson(), data = corales)
summary(glm12)
fit.boot <- boot.stepAIC(glm12, data = corales, B = 100) # That's it !
fit.boot

```

Los Porifera y las algas de tipo: 
2     - Branching.erect.algae  
3   - Coarsely.branched.algae  
4                - turf.algae  
5 - Branching.coralline.algae  
6   - Jointed.calcareus.algae 
  
No son buenas predictoras del la abundancia (reclutamiento) de los corales!

Sheet.like.algae + Filamentous.algae + 
    Crustose.algae + Thick.leathery.algae + unbranched.erect.algae
    SI son buenas predictoras




```{r sesion}

sessionInfo()



```

